### 2021-03-19 paper/foss4g/foss4g-zonebuilder-abstract.Rmd
Add abstract for #33

---
output: github_document
---

<!-- Brief: https://callforpapers.2021.foss4g.org/foss4g-2021-academic/ -->

## Zonebuilders: cross-platform and language-agnostic tools for generating zoning systems for urban analysis and modelling

Zones are key building blocks of analysis and models (mental and statistical) of urban and ecological systems.
Used in a range of fields from biodiversity assessment to transport planning, spatially contiguous areal units that break-up continuous space into discrete chunks are a foundation on which many approaches to geographical research build.
Techniques including origin-destination analysis, geographically weighted regression, and choropleth mapping rely on zones with appropriate sizes, shapes and coverage.
However, open access administrative boundaries representing locally specific zoning systems are either inappropriate or non-existent in many places, for a number of reasons. 
Administrative zoning systems are often based on arbitrary factors; they are generally inconsistent between different cities/regions; and their highly variable sizes and shapes make them hard to analyse.

In this paper we propose a scalable solution to these problems: tools that can auto-generate zones based on minimal input data.
We demonstrate a particular implementation of the approach --- the ClockBoard zoning system, which consists of 12 segments divided by concentric rings of increasing distance, creating a consistent visual frame of reference for cities that is reminiscent of a clock and a dartboard --- and propose alternative zoning systems.
Vitally from software freedom, transparency and accessibility perspectives, we propose cross-platform and language agnostic implementations to enable as many people as possible to generate bespoke zoning systems for their own purposes.
We present work in progress towards a framework enabling cross-platform and language-agnostic generation of zoning systems, with a core library written in Rust with small and fast binaries available for all major operating systems.
Higher level implementations are provided through an R package (published on the Comprehensive R Archive Network, CRAN), a Python Package, a QGIS plugin (work in progress) and web user interface.
We conclude by discussing future directions of travel in zoning system development and the possibilities of publishing new geographical methods for use in multiple free and open source frameworks by using low level and cross-platform implementations such as Rust as the foundation.