### 2021-04-06 paper-latex/zonebuilder-paper.Rmd

---
title: "Habits"
output:
  pdf_document:
    template: josis_template.tex
---

test

```{r, eval=FALSE, echo=FALSE}
# https://bookdown.org/yihui/rmarkdown/pdf-document.html#custom-templates-1
tinytex::install_tinytex()
tinytex::pdflatex("~/paper-latex/josis_template.tex")
```